\OurTool{}: Integrative Probing Analysis of Nucleic Acids Empowered by Multiple Accessibility Profiles\\[1em] Supplementary material

### 2019-12-22 Response-NAR-Round1.tex

	\Comment{First, Rsample’s sampling routine is quite different than IPANEMAP’s and could be a relevant factor. It entails sampling from SHAPE reactivity distributions followed by partition function optimization. The clustering’s contribution to performance gains can be assessed by applying such clustering to Rsample’s output sample, right before the MEA step. Such analysis is useful because if this routine is key to improvements, it may be adopted by numerous other methods that analyze Boltzmann samples.}

	\Answer{In the case of single-state RNAs, highlighting the shared sampling/clustering paradigm with \Software{RSample} turned out to be misguided on our part, since \Software{RSample} simply computes the MEA of the whole ensemble rather than to rely on a sampling/clustering combination. We have modified our discussion on the comparison with \Software{RSample} to reflect the fact that clustering is simply not used by \Software{RSample} in this setting.}

@@

	\Comment{Second, my understanding is that a sample size of 1,000 was used in all benchmarks. However, unlike the 1,000 size-convention, Spasic et al. used 10,000 structures in their benchmarks. It is possible that this deeper sampling is necessary to Rsample’s performance. I think it would be fair to compare IPANEMAP and Rsample performances using 10,000-structure samples. }

	\Answer{We agree that our description may have been imprecise in that respect, but we did use the recommended numbers of samples in our execution of \Software{RSample}. More precisely, as described in \cite{Spasic2017}, two different rounds of sampling were used in the method: a) to estimate the corrective terms; b) to identify alternative conformations, if several conformation are suspected. The recommendations of \cite{Spasic2017} are to use 10,000 samples for (a), and 1,000 samples for (b), which was done for the reported predictions. In the submitted manuscript we only referred to (b) for the number of samples, which was not very useful to the reader since step b) is never performed by \Software{RSample} on the benchmark.}

	\Comment{Third, the authors mention that Rsample was “shown to perform favorably against a comprehensive collection of state-of-the-art methods…” This is incorrect. In fact, based on a t-test, \cite{Spasic2017} concluded that “all methods that use experimental restraints perform similarly.” A similar conclusion, also based on a statistical test, was reported by \cite{Deng2016}.}

	\Answer{We thank the reviewer for pointing out this remark in the \cite{Spasic2017} paper which we overlooked in our initial reading. We corrected the corresponding sentence in our revised manuscript to
		\Quote{It was shown to perform \Rev{comparably as} a comprehensive collection of state-of-the-art methods in probing-guided structure prediction, including \Software{RME}~\citep{Wu2015}, \Software{RNAprob}~\citep{Deng2016}, \Software{RNAprobing}~\citep{Washietl2012}, \Software{RNAsc}~\citep{Zarringhalam2012} and the \Software{fold} utility from the \Software{RNAStructure} suite~\citep{Reuter2010}.}
		 This does not however challenge the conclusions drawn from this section, as \Software{RSample} remains as good a representative of the current state-of-the-art for mono-probing structure prediction.}

	\Comment{This matters because it points out a major weakness of this work, namely, it does not consider the statistical significance of its findings despite reporting substantial variability between predictions. Additionally, it highlights the possibility that a comparison to another method may result in different findings since there are a few sequences with dramatic performance gaps, which were not considered by \cite{Spasic2017}}

	\Answer{The significance (or lack thereof) of claims in the context of RNA benchmarks is, of course, a legitimate concern, and we thank the reviewer for pointing it out. The way it is used in \cite{Spasic2017} should probably be taken with a grain of salt due to the dubious obedience of produced data to classic assumptions (\emph{eg} normality of distribution), but it nevertheless represents the best option available at the time.

	In the context of comparing the performances of \OurTool{} to \Software{RSample}, we performed a two-tailed Student $t$-test on paired data with unequal variance. The test rejects the $H_0$ hypothesis \emph{Both methods have similar average performance}, with P-value of $6\times 10^{-3}$, much lower than the usual significance threshold $\alpha=5 \times 10^{-2}$  (computed statistics $t=2.98 > t_{\text{crit}}=2.07$).
	}

@@

	\Comment{Fourth, I think the stability of predictions over 10 independent samples should have been compared between the two algorithms. I could find stability results for IPANEMAP only (Table S5). In my experience, reproducibility of predictions is the most challenging issue with Boltzmann-based algorithms, especially when relying on clustering techniques. }

	\Answer{As observed above, on the \cite{Hajdin2013} dataset, \Software{RSample} always ends up running an MEA algorithm, an approach which is deterministic for a given set of base pair probabilities. Moreover, the base pair probabilities are based on correcting terms that are themselves estimated from 10~000 sampled structures (a number chosen to ensure stable estimates). Therefore, in theory, \Software{RSample} is virtually deterministic in its predictions, and quite unsurprisingly we could not detect a single variation of its predictions within 10 executions over the \cite{Hajdin2013} dataset.}

@@

	\Comment{Fifth, the Methods section indicates IPANEMAP does not use default “m” and “b” values for Deigan’s algorithm because different values seem to work better. It is unclear to me which of the analyzed datasets was used to determine these values or how the “m” and “b” parameters were optimized. These details should be included in the text. Moreover, a comparison between the algorithms should train them on the same dataset. If the Hajdin dataset was used in IPANEMAP’s parameter optimization, I think it is fair to re-optimize Rsample’s parameters too. This is because of the few sequences that were not considered by Spastic et al., for which performance gains are huge. The different training sets could be a factor in the reported performance gains. }

	\Answer{The \cite{Hajdin2013} was not used in the initial training.}

@@

	\Comment{Sixth, according to the abstract, IPANEMAP includes an unsupervised learning component. However, if the Deigan’s “m” and “b” parameters were optimized against known structures, then I don’t see which parts are unsupervised. This is another reason why I request to clarify the parameter optimization step. }

	\Answer{
		The distance-based clustering of structures, arguably the most crucial component of \OurTool, is fully unsupervised. We do not believe that a loose calibration of two numerical parameters should raise a reasonable concerns of overfitting.
	}

@@

	\Comment{The bulk of the manuscript describes various benchmarks, typically comparing average accuracies over multiple IPANEMAP runs, over multiple RNAs, or over multiple conditions. However, a comparison of two sets of data points should consider their means in the context of their std’s, otherwise it is often not too informative or conclusive. What I’m missing here are classical tests that quantify the statistical significance associated with each comparison. This is missing in all tests conducted in this work, and I suspect that a few of the reported mean differences might not be very significant. One example of where a test is needed is the comparison to Rsample -  see my comment above. Another example is the comparison to MFE and MEA (Table 2). There are several more. Note that if multiple tests are jointly considered/reported then p-value adjustment (e.g., B-H) might be warranted.}

	\Answer{
	Despite our previously-stated objections, we understand that such metrics are expected for new methods. We assessed the statistical significance of our main assertions using the two-tailed Student $t$-test on paired data with unequal variance, as recommended by \cite{Xu2012}:
	\begin{itemize}
		\item \cite{Hajdin2013} dataset.\\
		Hypothesis $H_0$: \OurTool and \Software{RSample} have similar expected GM\\
		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=6\times 10^{-3}$ $\to$ Rejected at $\alpha=5\times10^{-2}$ 
		\item \cite{Cordero2012} dataset.\\
		  Due to the limited size of the dataset (6 RNAs), none of our observations regarding the average behavior are sufficient to reach statistical significance. We have toned down the title of the corresponding section, and added a remark to the conclusion of the associated results section.
		\item \didy{}  dataset.\\
		Hypothesis $H_0$: \OurTool and MFE have similar expected MCC\\
		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=2.2\times 10^{-2}$ $\to$ Borderline rejected at $\alpha=5\times10^{-2}$\\
		Hypothesis $H_0$: \OurTool and MEA have similar expected MCC\\
		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=14.4\times 10^{-2}$ $\to$ Not significant at $\alpha=5\times10^{-2}$ \\
		Hypothesis $H_0$: Joint analysis of triplet have same MCC as average of three member conditions\\
		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=1.4\times 10^{-9}$ $\to$ Rejected at $\alpha=5\times10^{-2}$ \\
		Hypothesis $H_0$: Joint analysis of triplets and pairs of conditions have same MCC\\
		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=2.1\times 10^{-2}$ $\to$ Borderline rejected at $\alpha=5\times10^{-2}$ \\
		\item \cite{Cheng2017} dataset.\\
		Hypothesis $H_0$: WT and WT+1M have similar expected MCC\\
		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=9.7\times10^{-4}$ $\to$ Rejected at $\alpha=5\times10^{-2}$\\ 
		Hypothesis $H_0$: WT+1M and WT+2MC have similar expected MCC\\
		\phantom{Hypothesis $H_0$: }$\mathbb{P}(H_0)=2.4\times10^{-2}$ $\to$ Rejected at $\alpha=5\times10^{-2}$
	\end{itemize}
	}

	\Comment{The authors generated 16 new datasets that enabled them to consider various ways of combining data from multiple conditions. This is impressive, with two caveats. First, only a single RNA of medium length was probed.}

	\Answer{
		It must be noted that we do not base our main conclusion (benefit from the joint analysis of multiple reactivity) solely on the \didy{} case study, but also on the Cordero dataset, where the best performances are also observed upon considering a combination of reactivity profiles associated with DMS, CMCT and NMIA probes. A similar trend is also observed on the \cite{Cheng2017} dataset.
	}

	\Comment{Second, it is my understanding  that no replicate data was generated or analyzed, with the exception of SHAPE-CE. Typically, methods are tested on at least a few RNAs of varying lengths and whenever possible on datasets like Hajdin et al.’s. From such studies, we know that data-directed gains vary substantially between RNAs. For this reason, I would be hesitant to draw any conclusions from probing data for a single RNA.

	That said, I do not expect the authors to generate more data. I do think it should be emphasized that this is merely a case study. The purpose of this study and what one could conclude from it should also be clarified. While the subsection’s title (page 7) indicates this is a case study, I believe that some statements in the Abstract and Discussion require further substantiation and should be toned down.}

	\Answer{We already clarified that this is merely a case study. We wish we could do more (e.g. test on a wider dataset) but this is simply unrealistic, and way beyond the standards enforced for the dissemination of a new {\bf method} in structural bioinformatics.

		 However, it must be noted that we do not base our main conclusion (complementarity of probing data, benefit of their integration) solely on the \didy{} case study, but also on the Cordero dataset, where the best performances are also observed upon considering a combination of reactivity profiles associated with DMS, CMCT and NMIA probes. 
	}

@@

	\Comment{The other reason I feel these results should be framed as a case study is that replicate data was not considered in the analysis. The manuscript attempts to differentiate between conditions or between algorithms, but the variation within each condition should be assessed in order to confirm that observed differences are in fact meaningful. }

	\Answer{
		\TODOBruno{Did we produce replicates?}
	}

	\Comment{The assessment of mono-probing predictions on page 8 does not seem to result in a clear and significant take-home message. Accuracies seem to be comparable between IPANEMAP, MEA, and MFE, but I don’t see what we learn from this. Note also that the data-directed MEA and MFE algorithms were optimized for SHAPE data only, hence they might not do as well on non-SHAPE data (Deng et al., RNA, 2016).}

	\Answer{We agree, and in fact do not draw any conclusion from these results. We only report those values as references, to support the hypothesis that the MCCs achieved by \OurTool are probably more representative of the reactivity profiles/condition than of a specific predictive model.}

	\Comment{Additionally, there is strong A/C bias in DMS data obtained by stop-based protocols. I don’t know if it was accounted for during normalization step, if normalization was indeed applied. }

	\Answer{
		We were unsure how to interpret this remark: \Ab{} and \Cb{} are indeed the only reactive nucleotides, so the normalization step is indeed restricted to positions featuring those two nucleotides. If the reviewer is referring to the observations of \cite{Sexton2017} of opposite {\em \Ab{} vs \Cb} biases for RT-based and mutations-based probing respectively, we did not account for it. Indeed, while including corrective terms would seem to represent an opportunity for improved predictions, we would rather treat of reactivity data homogeneously than achieve the best MCCs possible. More details in our response to Comment \ref{arg:homogenous} of Reviewer \#3.
	}

	\Comment{I also don’t know which datasets were used to optimize IPANEMAP’s parameters, but as I commented above, a fair comparison should train all methods on the same data.}

	\Answer{Again, we object to such an adversarial view of this particular analysis (and, more generally, our work). \OurTool offers new capabilities (joint integration of multiple probing data), which were not supported by previous tools. Our extensive comparison to earlier efforts on mono-probing data is mainly meant to validate our specific implementation of the \cite{Ding2003} sampling/clustering in the context of single probing data.
	}

@@

	\Comment{Finally, I don’t see why IPANEMAP was compared to MEA and MFE in this subsection but to Rsample in the previous one. Any special reason why Rsample is not included? The authors may also want to consider moving this part to the Supp.}

	\Answer{
		Again, we include those MCCs for reference only, and leave their interpretation to the reader. We do not feel the need to include a huge collection of tools, as this is not a benchmark.
	}

@@

	\Comment{Please edit for typos and grammar. There are also frequent transitions between present and past tense within paragraphs, which don’t read well.}

	\Answer{In this revised version, we tried to enforce a consistent use of the past tense in the results section.}

@@

	\Comment{I think an explicit discussion of IPANEMAP’s limitations is warranted in Discussion or as an independent subsection.}

	\Answer{Although we already described some of those limitations in the method, we now include such a discussion more explicitly in our conclusion/discussion.
	\Quote{Current limitations of \OurTool include the lack of an explicit support for pseudoknots, which will be addressed in a future version, using  polynomial-time dynamic programming~\citep{Janssen2015}, or a iterative heuristic alternative~\citep{Hajdin2013}. The inherently stochastic nature of the sampling scheme may also be challenging for certain types of analysis. However, despite its stochastic foundations, \OurTool{} is typically stable in its predictions, as can be seen in Figure~\ref{fig:reproducibility} for all probing datasets produced over our \didy case study.
			Interestingly, the most notable exception to the general stability is observed for the probing-free prediction, consistent with the widespread interpretation of SHAPE-induced pseudo-potentials as focusing predictions onto a subset of the Boltzmann ensemble. A similar behavior is observed for the other datasets considered in our study (see Supp. Tables S2 and S5). %In the interest of reproducibility, ipanemap reports, and accepts as an option, the random seed used both within the main software and its non-deterministic dependencies, leading to identical repeats of the experiments.
			Finally, the current clustering method induce a growth of time and memory requirements that scales quadratically with the number of samples. While this allows a joint analysis of a dozen of reactivity profiles in less than an hour on a personal computer, we expect the ability to push the number of samples will lead to better, even more reproducible, predictions. To that purpose, we will explore embedding techniques and linear-time community detection algorithms as alternatives to MBkM~\citep{Sculley2010}.
	}	
}

@@

	\Comment{Many details could be moved to the Supp. to improve readability, specifically descriptions of results that essentially parse a relevant table for the reader. Often, the text gets so focused on the trees that one can’t see the forest. Briefly stating the main trend or conclusion with reference to a table should suffice. Examples include the last 2 paragraphs on page 9, first paragraph on page 10, and third paragraph on right column on page 11 (“For the WT…”).}

	\Answer{We respectfully disagree with this proposal of an alternative writing style, and prefer to include sufficient factual statements to let the readers form an opinion regarding our reasoning and conclusion (at the risk of being tediously descriptive at times). }

@@

	\Comment{Much space is allocated to outlier detection (cluster B) and then to describing analyses with and without cluster B. I feel that the authors can safely designate cluster B as an outlier and exclude it from the analyses described in the main text. Analyses with cluster B can be described in the Supp.}

	\Answer{Again, we respectfully disagree with this vision. What is important about cluster B is not only the fact that it can be identified from our new metrics (more conclusively than than from the raw reactivities), but also that it can be used to assess the impact of contamination by an erroneous/outlier condition.}

@@

	\Comment{Were all the analyzed datasets normalized by any standard method (e.g., 2\%-8\% or box-plot)? I noticed that the mutate-and-map dataset was renormalized, and to the best of my knowledge, the Hajdin dataset is also normalized. However, I didn’t see any information re: Cordero dataset, and more importantly re: the newly generated data analyzed here. Normalization is especially important in the context of differential analysis between conditions.}

	\Answer{\TODOAfaf{Can you confirm that a similar normalization was used?}}

@@

	\Comment{Fig. 5: Why was k set to 8 as opposed to, say, 4 or 5? As the authors note, determining the number of clusters is tricky and can significantly impact results. }

	\Answer{We came to this conclusion from a visual inspection of the distance matrix in Figure 5, reproduced below. 
		\Quote{
	\includegraphics[width=.45\textwidth]{graphs/didy/PCA} \hfill
	\includegraphics[width=.45\textwidth]{graphs/didy/bi_clustering}\\
	}

		This number may not seem to be supported by a inspection Figure 5 (PCA, left above), but it must be reminded that a PCA is simply a projection. Points that are distant in the PCA are typically distant in the original higher-dimensional space, but close points in the PCA may also be quite distant due to the missing depth perception.

		On Figure 5, we already see a clear separation between B, D, C, A, and H. Moreover, merging clusters E, F and G would lead to the aggregation of substantially different conditions, as can be seen in Figure 6 (right above), so we considered 8 conditions.}

@@

	\Comment{The authors compare stop-based to mutation-based experiments. However, the term “stop-based” is somewhat misleading because there is another variable here - the modification assay platform. MaP protocols must be assayed by NGS whereas the stop-based protocols used in this work rely on CE for modification detection. It is possible that the differences observed between MaP and CE data (see e.g., Fig. 5) have little to do with the stop strategy and more with the sequencing platform (CE vs NGS). The use of “stop-based” might create an impression that this feature is necessarily responsible for these differences. The ideal comparison should be between experiments such as SHAPE-MaP and SHAPE-seq, but given that experiments were already performed, I recommend changing the term and clarifying the technical differences between these subsets of conditions.}

	\Answer{We agree that SHAPE CE should not be mischaracterized as a representative for all SHAPE protocols based on RT stops (notoriously including SHAPE-Seq). Our description was misleading in that respect, so we modified and removed the corresponding column in Table 2.}

### 2019-12-22 Response-NAR-Round1.tex

	\Comment{Does IPANEMAP perform better when sampling deeper? More generally, was sample size of 1,000 selected following performance optimization or simply because there is a convention in the field that 1,000 suffices for reproducibility at the base-pair level?}

	\Answer{
		We were happy with the overall reproducibility of our results with 1~000 structures, which has been the standard practice in the field since the original \cite{Ding2003} paper. Moreover, our computational requirements scale quadratically with the number of structure, so we recommend to stay with this number, which represents in our opinion a good tradeoff between reproducibility and efficiency.

		To address this question more precisely, we reexecuted the \didy benchmark using a sample size of 2~000, and observed a minor non-significant variation in the MCC (ie the results did not allow to refute the hypothesis of equal expected MCCs for both experiments).}

@@

	\Comment{Table 1 could be moved to the Supp.}

	\Answer{We disagree and, while anecdotal, we believe this figure illustrates nicely a case where the addition of data leads to an improvement of the model.}

@@

	\Comment{Table 2 summarizes performance differences at the Mut/Stop and -/+Mg levels. I wonder if such differences could have been detected already at the reactivity level and without the need for Boltzmann sampling. For example, one could lump all “Mut” profiles together as replicates for condition A and all “Stop” profiles as replicates for condition B and then perform differential reactivity analysis to spot differentially reactive regions and their statistical significance.}

	\Answer{For this revision, we added a principled comparison of the raw reactivity profiles based on correlation of reactivity. Its results are now included in the supplementary material and reproduced below
	\Quote{\centering\includegraphics[width=.64\textwidth]{graphs/didy/reactivity_correlation.pdf}}
	As can be seen in this matrix, "MaP" conditions essentially lump in two classes at the reactivity level (that do not fully align with the "MaP" clusters A/C/D). However, "CE" conditions do not form a clear cluster, and in many case correlate better with a "MaP" conditions than with "CE" conditions in the same cluster.
}

@@

	\Comment{In Table 2, the two conditions in cluster C perform quite differently: 74\% and 85\%. Can the authors suggest why this happens? Can the performance differences be picked up by a differential reactivity analysis? For example, there may be a certain differentially reactive region, which the ensemble distance measure does not pick up. }

	\Answer{Indeed, the difference can be traced back, for the conditions in C, to a reactivity in the 5' terminal region that is greater for  \OneMSevILUThreeMg than for \OneMSevILUMg, resulting in a loss of the basal helix in the model predicted by \OneMSevILUThreeMg. This variation, coupled with the shift of a small helix, is alone responsible for the change of performances (the MCC does not take prisoners...). }

@@

	\Comment{Does bi-probing improve precision (std) over mono-probing? Similarly, does probing with multiple similar conditions improve precision compared to mono- or bi-probing (Table 3)?}

	\Answer{As stated in the manuscript, bi-probing does not improve the standard deviation of the MCC (8\% for mono-probing, 8.4\% for bi-probing). Considering the sets of similar conditions in Table 3 increases the stdev to around 10\%. However, this observation merely stems from a statistical artifact, as the large number of conditions in Cluster B leads to a multiplication of MCC around 60\%, resulting in what would appear to be a bimodal distribution ($\to$ large variance/stdev) without reflecting any underlying reality.}

	\Comment{Performance assessments in this work focus primarily on averages although precision is an important factor, especially when statistical sampling is involved and large variation in prediction accuracy is observed for different RNAs. The conclusion that one might not gain much or at all from increasing the no. of probes or using similar ones overlooks the notion that redundancy potentially increases our confidence in the results. For example, replicates are inherently redundant and ideally not too diverse, yet they are generated to decrease std around the mean although the mean is always lower than some of the point estimates. A related question is as follows.

	For bi-probing, if we repeat predictions with 10 independent Boltzmann samples, which strategy is more precise: averaging predictions between the two probes or the corresponding bi-probing prediction?}

	\Answer{
		The meaning of \emph{averaging the predictions} is unclear to us in this context (our method produces 2D model(s), how does one average models?), and could be interpreted in several ways that we try to cover in the following sentences.

		If the intent is to compare the MCC of a bi-probing prediction with the average of the two mono predictions, this is precisely the content of Figure 7.B, and our analysis does not reveal a substantial reduction of the stdev.

		The intent may also be to compare, for a given pair of conditions, the stdev of the MCC to the stdev of the mean MCC over two independent mono executions (by running several experiments). In this setting, we either analyze a single pair, and obtain results that are probably not significant, or we need to report summary statistics over stdevs of a dataset of pairs of conditions, something that would not only be heavily time-consuming, but also hard to interpret. We will welcome clarification over this point over the next round of reviews.

		However, if the objective metric of quality is the reproducibility of predictions using \OurTool (\emph{i.e.} to decrease the stdev as low as possible), a very simple solution is to perform 10+ independent runs of \OurTool, and elect the dominant conformation. Indeed, in most cases, \OurTool produces a well-defined structure with high probability, as suggested by Figure 8 (WT) and 9. By running \OurTool repeatedly, the frequency of the dominant structure becomes an estimator, with variance decreasing on the number of iterations, for the emission probability of the structure. This process ultimately converges towards a deterministic election of the most probable structure, and should already be fairly close to determinism after 10 iterations. Moreover, the time requirement scales linearly with the number of iterations, as opposed to quadratically if one were to multiply the number of samples.

		In the future, we also wish to report, and letting the user specify, a \Def{random seed}, which will  allow the exact reproduction of an analysis. 
		However, such a feature is currently not supported by \Software{RNAsubopt}, so we must first work towards its inclusion to the \Software{ViennaRNA} package.
		}

@@

	\Comment{Table 5: I do not think it demonstrates a significant finding. All subsets include $1M7-MaP_{IL}^{MG}$, whose mono performance is already 0.85, as compared to 0.853. According to Table 2, it is the best-performing condition. My guess is that the 0.003 gain is an artifact rather than a manifestation of complementary information embedded in the additional conditions. If the point made is that it’s advantageous to jointly analyze conditions rather than average them out, then I think it could be made in the context of the other results. However, as shown here, using too many conditions might negatively affect performance, so not sure if robust guidelines can be derived from these results.}

	\Answer{Upon further inspection of the prediction, it appears that the reviewer's intuition is correct and the improvement in comparison to the best of the group is marginal. We have removed the Table, and any mention to those specific triplets in the text.}

@@

	\Comment{Fig. 9: $1M7-MaP_{IL}^{MG}$ is a top performer on the Lariat data per Table 2, but its MCC variance over 10 runs is high per Fig. 9. If I interpret Fig. 9 correctly, the 85\% MCC from Table 2 is not shown for this condition and maybe resulted from another IPANEMAP run. In any case, I wonder if Table 2 lists the results of single runs or averages over 10 runs. If single runs are listed, how representative are they for the two conditions with high variances (Fig. 9)? Maybe replace them with average runs? Same questions for results with multiple probes. Note that the case for $NMIA-MaP_{IT}$ is similar as 80\% seems to be max performance.}

	\Answer{
		We would like to sincerely thank this reviewer for noticing and voicing this apparent inconsistency, which led us to realize a severe mistake in Figure 9. For reasons to be determined, the labeling of the y-axis (MCC) ended up being dilated, showing MCCs in the 70\%/82\% range. The revised version of Figure 9, reproduced below, now features corrected axis and is now consistent with the averages MCC reported in Table 2.
		\Quote{
				{\centering
					\begin{minipage}{.58\linewidth}
					\includegraphics[width=\linewidth]{graphs/boxplotreproducibility}
					\end{minipage}\hfill
					\begin{minipage}{.4\linewidth}
	\Heatset{min=60,   
		max=83,   
		max colour=Aquamarine3, % colour at maximum
		min colour=BadCol,      % colour at minimum
		Min colour=BadCol, % colour for values below min
		Max colour=Aquamarine3   % colour for values above max
	}
					\newcommand{\M}[1]{\multicolumn{1}{l}{#1}}
						\begin{adjustbox}{max width=\linewidth}
							\begin{tabular}{@{}lccHHH}
								\toprule
								&          &           & \multicolumn{3}{@{}c@{}}{MCC\%}              \\
								Condition                   &  Clust.  & Mg$^{2+}$ & \M{\relsize{-1}\OurTool} & \M{MFE} & \M{MEA} \\ \midrule
								\OneMSevILUMg               & \Bull{C} &    \Mg    & 85                       & 82      & 82      \\
								\OneMSevILU                 & \Bull{A} &   \NoMg   & 84                       & 82      & 84      \\
								\NMIAMg                     & \Bull{A} &    \Mg    & 81                       & 80      & 80      \\
								\NMIA                       & \Bull{A} &   \NoMg   & 80                       & 69      & 69      \\
								\OneMSevILUThreeMg          & \Bull{C} &    \Mg    & 74                       & 75      & 75      \\
								\NMIAMgCE                   & \Bull{E} &    \Mg    & 73                       & 72      & 73      \\
								\NAIMg                      & \Bull{G} &    \Mg    & 73                       & 68      & 70      \\
								\BzCNMg                     & \Bull{H} &    \Mg    & 71                       & 68      & 54      \\
								\OneMSevMgCE                & \Bull{E} &    \Mg    & 70                       & 69      & 70      \\
								\CMCTMg                     & \Bull{G} &    \Mg    & 70                       & 68      & 73      \\
								\OneMSevILUThree            & \Bull{D} &   \NoMg   & 64                       & 56      & 61      \\
								\DMSMg                      & \Bull{F} &    \Mg    & 63                       & 66      & 67      \\
								\NMIACE                     & \Bull{B} &   \NoMg   & 61                       & 60      & 60      \\
								\OneMSevCE                  & \Bull{B} &   \NoMg   & 60                       & 61      & 61      \\
								\BzCN                       & \Bull{B} &   \NoMg   & 60                       & 59      & 59      \\
								\NAICE                      & \Bull{B} &   \NoMg   & 60                       & 59      & 59      \\ \midrule
								Avg Technology              &          &  $-$  & 78                       & 74      & 75      \\
								&          &  $-$  & 66                       & 65      & 65      \\
								Avg -/+ Mg$^{2+}$           &          &   \NoMg   & 67                       & 64      & 65      \\
								&          &    \Mg    & 73                       & 72      & 71.5    \\ \midrule
								{\bfseries Average overall} &          &  $-$  & 70.5                     & 68      & 68.5    \\ \bottomrule
								&          &           &                          &
							\end{tabular}
					\end{adjustbox}
					\end{minipage}
					\\}
		}

		Average MCC values over 10 runs are indeed reported in Table 2, and an independent dataset was generated to create Figure 9, leading to minor discrepancies. For multiple probes, only a single result was generated.
	}

@@

	\Comment{Page 8, line 55: 1000 samples or 1000 structures per condition?}

	\Answer{Fixed.}

	\Comment{Methods, page 3: The definition of highly similar clusters uses a default value $\delta=1$. My understanding is that this means that highly similar clusters must have the exact same centroids, such that minor variation is not allowed. Is that correct?}

	\Answer{Two centroids distant by at most one base-pair (included) will be deemed highly similar, so a (very) minor variation will be tolerated.}

@@

	\Comment{Methods, page 3: Is $E_{d}(S)$ the pseudo-energy term of the total score or is it the entire score, including the NNTM free-energy? }

	\Answer{$E_{d}(S)$ denotes the entire score, including NNTM free energy plus the energy bonus.}

@@

	\Comment{My impression is that the main purpose of the clustering was to detect the outliers in cluster B. Since I find this to be a technical piece of the entire analysis, and not a very interesting one, I suggest moving most of the text + Fig. 5 \& 6 to the Supp.}

	\Answer{We already addressed this comment above, and respectfully disagree with this assessment.}

@@

	\Comment{Page 5, benchmarks on Cordero et al.’s dataset: It is noted that the prediction is the centroid of the largest probability cluster. My understanding, based on the method’s section, is that the prediction should be the centroid of the pareto-optimal clusters. It is unclear whether this is a modification of the method, and if yes, if it applies to all other benchmarks or not. Please clarify. }

	\Answer{In a mono-probing setting, due to properties of the two criteria used for the multi-objective Pareto optimization, it is rigorously impossible for \OurTool{} to return more than one cluster.  Indeed, the cluster $C^\star$ having max stability $s^\star$ is such that: either $C$ has support $t^\star=1$, in which case it dominates any other cluster with signature $(s<s^\star,t\in\{0,1\})$; or $C$ has support $t^\star=0$, ruling out the possibility of $t=1$ for any other cluster (since it would imply $s>s^\star$). Note that \OurTool also reports dominated clusters for furter inspection.
	}
